LME-MPM applied to quasi-brittle fracture.

### 2020-05-21 _region_.tex
Cambios 20-5-2020 (II) : Borrador terminado

This problem here presented is devoted to compare how much \acrshort{lme}
approximants are able to improve the result \textit{versus} standard
linear interpolation. It consists in a square plate of size $H = 1$ containing
an initial edge crack of length $0.25 \cdot H$ loaded in a pure mode I
by displacement control on the outer flanks of the plate, Figure
\ref{fig:geometry-cracked-panel-mode-I}.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  \centering
  \includegraphics[width=0.4\textwidth]{Figure-Mode_I}
  \caption{Geometry and boundary condition of the drop-weight impact test.}
  \label{fig:geometry-cracked-panel-mode-I}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The constitutive model consider for numerical experiment is a
linear-elastic Hookean material, where Young's modulus \gls{E} = 1.06, the
Poisson's ratio \gls{nu} = 0.333, and critical energy-release rate
\gls{Gf} = 0.0001.
This set of parameters was taken from \cite{Pandolfi_2012}. In
\acrshort{mpm}, two different discretization are required. In one hand
a cartesian grid of nodes is consider with a nodal spacing value of
0.025. On the other hand, the plate will be modeled with a initial
layout of four particles per element and occupying the Gauss quadrature
positions.

Figure \ref{fig:Reactions-cracked-panel-mode-I} clearly shows the
presence of wiggles in both loading and fracture stages
in the reaction-displacements curve when linear interpolation technique is employed. In
contrast, \acrshort{lme} simulation does not exhibit these spurious
oscillations and remains linear. Besides that, linear interpolation
produces stiffer results than \acrshort{lme}. This can be attributed
to imprecision in the stress field due to grid-crossing phenomena.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{Figure-Reactions-Mode-I}
  \caption{Evolution of the reaction forces plotted \textit{versus}
    the imposed displacement for linear interpolation technique and
    the \acrshort{lme} approximants.}
  \label{fig:Reactions-cracked-panel-mode-I}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
On the other hand, Figure \ref{fig:Stress-cracked-panel-mode-I}
exhibits more clearly the difference between linear interpolation and
\acrshort{lme}, since numerical instability becomes rather significant
when particles crosses the boundary of the element in the linear
case. Although this behavior does not affect significantly to the peak
stress value in a hookean material \cite{Zhang_EE_2020}, if
hyperelastic materials are consider for future simulations,
inaccuracies in the stress field could affect dramatically to the
final result. On the other hand \acrshort{lme} produces soft stress
field evolution even after breaking    
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\centering
\subfigure[t = 25]{\includegraphics[width=0.73\textwidth]{./Figure-Stress-0100-mode-I.png}}
\subfigure[t = 60]{\includegraphics[width=0.73\textwidth]{./Figure-Stress-0250-mode-I.png}}
\subfigure[t = 100]{\includegraphics[width=0.73\textwidth]{./Figure-Stress-0398-mode-I.png}}
\caption{Evolution of the stress tensor magnitude for linear
  interpolation (pictures in the left side), and \acrshort{lme}
  (pictures in the right side). Both simulations performed with an
  eigenerosion algorithm.}
\label{fig:Stress-cracked-panel-mode-I}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
An important consideration regarding the presence of oscillations once
both parts of the panel are separated, Figure
\ref{fig:Reactions-cracked-panel-mode-I}. These phenomena should not be
attributed to the eigenerosion algorithm since the fracture process is
over. They are due to the dynamic nature of the solver.

@@

Once the accuracy of the \acrshort{lme} approximants have been proven
in Section \ref{sec:3.1}, henceforth will be adopted without comparing
with linear interpolation. Therefore this section is devoted to proof
the accuracy of the eigensoftening algorithm to simulate the behavior
of quasi-brittle materials under dynamic loading. An example of this
load case is the three-point bending test on a notched beam is
conducted under impact loading. This kinds of experiments are
challenging to reproduce numerically with an explicit solver since
fracture occurs in a period of milliseconds and high strength
materials are involved. Accordingly, really small time steps are
required to achieve a simulation numerically stable.

@@

And finally horizontal stress distributions and crack propagation are
extracted, see Figure \ref{fig:Stress-vs-damage-impact-test}. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\centering
\subfigure[t = 0.27 ms]{\includegraphics[width=0.83\textwidth]{./Figure-impact-test-Damage-vs-Stress-0060}}
\subfigure[t = 0.36 ms]{\includegraphics[width=0.83\textwidth]{./Figure-impact-test-Damage-vs-Stress-0080}}
\subfigure[t = 0.45 ms]{\includegraphics[width=0.83\textwidth]{./Figure-impact-test-Damage-vs-Stress-0100}}
\caption{Evolution of the stress tensor magnitude for linear
  interpolation (pictures in the left side), and \acrshort{lme}
  (pictures in the right side). Both simulations performed with an
  eigenerosion algorithm.}
\label{fig:Stress-vs-damage-impact-test}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\end{document}
\message{ !name(2020_EFM_MPM_Eigensoftening.tex) !offset(-1010) }

>>>>

This problem here presented is devoted to compare how much \acrshort{lme}
approximants are able to improve the result \textit{versus} standard
linear interpolation. It consists in a square plate of size $H = 1$ containing
an initial edge crack of length $0.25 \cdot H$ loaded in a pure mode I
by displacement control on the outer flanks of the plate, Figure
\ref{fig:geometry-cracked-panel-mode-I}.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  \centering
  \includegraphics[width=0.4\textwidth]{Figure-Mode_I}
  \caption{Geometry and boundary condition of the drop-weight impact test.}
  \label{fig:geometry-cracked-panel-mode-I}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The constitutive model consider for numerical experiment is a
linear-elastic Hookean material, where Young's modulus \gls{E} = 1.06, the
Poisson's ratio \gls{nu} = 0.333, and critical energy-release rate
\gls{Gf} = 0.0001. This set of parameters was taken from \cite{Pandolfi_2012}. In
\acrshort{mpm}, two different discretization are required. In one hand
a cartesian grid of nodes is consider with a nodal spacing value of
0.025. On the other hand, the plate will be modeled with a initial
layout of four particles per element and occupying the Gauss quadrature
positions.\\

Figure \ref{fig:Reactions-cracked-panel-mode-I} clearly shows the
presence of wiggles in both loading and fracture stages
in the reaction-displacements curve when linear interpolation technique is employed. In
contrast, \acrshort{lme} simulation does not exhibit these spurious
oscillations and remains linear. Besides that, linear interpolation
produces stiffer results than \acrshort{lme}. This can be attributed
to imprecision in the stress field due to grid-crossing phenomena.\\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{Figure-Reactions-Mode-I}
  \caption{Evolution of the reaction forces plotted \textit{versus}
    the imposed displacement for linear interpolation technique and
    the \acrshort{lme} approximants.}
  \label{fig:Reactions-cracked-panel-mode-I}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
On the other hand, Figure \ref{fig:Stress-cracked-panel-mode-I}
exhibits more clearly the difference between linear interpolation and
\acrshort{lme}, since numerical instability becomes rather significant
when particles crosses the boundary of the element in the linear
case. Although this behavior does not affect significantly to the peak
stress value in a hookean material \cite{Zhang_EE_2020}, if
hyperelastic materials are consider for future simulations,
inaccuracies in the stress field could affect dramatically to the
final result. On the other hand \acrshort{lme} produces soft stress
field evolution even after breaking.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\centering
\subfigure[t = 25]{\includegraphics[width=0.73\textwidth]{./Figure-Stress-0100-mode-I.png}}
\subfigure[t = 60]{\includegraphics[width=0.73\textwidth]{./Figure-Stress-0250-mode-I.png}}
\subfigure[t = 100]{\includegraphics[width=0.73\textwidth]{./Figure-Stress-0398-mode-I.png}}
\caption{Evolution of the stress tensor magnitude for linear
  interpolation (pictures in the left side), and \acrshort{lme}
  (pictures in the right side). Both simulations performed with an
  eigenerosion algorithm.}
\label{fig:Stress-cracked-panel-mode-I}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
An important consideration regarding the presence of oscillations once
both parts of the panel are separated, Figure
\ref{fig:Reactions-cracked-panel-mode-I}. These phenomena should not be
attributed to the eigenerosion algorithm since the fracture process is
over. They are due to the dynamic nature of the solver.

@@

Once the accuracy of the \acrshort{lme} approximants have been proven
in Section \ref{sec:3.1}, henceforth will be adopted without comparing
with linear interpolation. Therefore this section is devoted to proof
the accuracy of the eigensoftening algorithm to simulate the behavior
of quasi-brittle materials under dynamic loading. An example of this
load case is the three-point bending test on a notched beam is
conducted under impact loading. This kinds of experiments are
challenging to reproduce numerically with an explicit solver since
fracture occurs in a period of milliseconds and high strength
materials are involved. Accordingly, really small time steps are
required to achieve a simulation numerically stable. Additionally,
this kind of simulations traditionally require the implementation of
contact algorithm to reproduce the hammer impact in the beam. In the
\acrshort{mpm}, the implementation of this feature can be bypassed to
obtain preliminary results. Nevertheless, this kind of algorithm could
be implemented to get ever more accurate solutions. Although this is
out of the scope of this document, future research of authors will overcome this
limitation.\\

@@

And finally horizontal stress distributions and crack propagation are
extracted, see Figure \ref{fig:Stress-vs-damage-impact-test}. Compared
with results obtained by \cite{Navas_2017_ES}, stress evolution

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\centering
\subfigure[t = 0.27 ms]{\includegraphics[width=0.83\textwidth]{./Figure-impact-test-Damage-vs-Stress-0060}}
\subfigure[t = 0.36 ms]{\includegraphics[width=0.83\textwidth]{./Figure-impact-test-Damage-vs-Stress-0080}}
\subfigure[t = 0.45 ms]{\includegraphics[width=0.83\textwidth]{./Figure-impact-test-Damage-vs-Stress-0100}}
\caption{Evolution of the stress tensor magnitude for linear
  interpolation (pictures in the left side), and \acrshort{lme}
  (pictures in the right side). Both simulations performed with an
  eigenerosion algorithm.}
\label{fig:Stress-vs-damage-impact-test}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\end{document}
\message{ !name(2020_EFM_MPM_Eigensoftening.tex) !offset(-1019) }
